### 2023-01-23 paper/admin/cover_letter.tex

\title{cover letter}
%
% See http://texblog.org/2013/11/11/latexs-alternative-letter-class-newlfm/
% and http://www.ctan.org/tex-archive/macros/latex/contrib/newlfm
% for more information.
%
\documentclass[11pt,stdletter,orderfromtodate,sigleft]{newlfm}
\usepackage{blindtext, xfrac, animate, hyperref, pxfonts, geometry}

  \setlength{\voffset}{0in}

\newlfmP{dateskipbefore=0pt}
\newlfmP{sigsize=20pt}
\newlfmP{sigskipbefore=10pt}

\newlfmP{Headlinewd=0pt,Footlinewd=0pt}

\newcommand{\journal}{Journal Name}
\newcommand{\articletype}{Article Type}
\newcommand{\title}{Paper title}
\newcommand{\corresponding}{Corresponding Author Name}

\namefrom{\vspace{-0.3in}\corresponding}
\addrfrom{
  Corresponding Author Institution\\
  Corresponding Author Department\\
  Corresponding Author Address\\
  City, State, Zip code}
\addrto{}
\dateset{\today}

\greetto{To the editors of \textit{\journal}:}

\closeline{Sincerely,}
%\usepackage{setspace}
%\linespread{0.85}
% How will your work make others in the field think differently and move the field forward?
% How does your work relate to the current literature on the topic?
% Who do you consider to be the most relevant audience for this work?
% Have you made clear in the letter what the work has and has not achieved?

\begin{document}
\begin{newlfm}
  We have enclosed our manuscript entitled \textit{\title} to be
  considered for publication as an \textit{\articletype}.  The
  manuscript... (insert one sentence summary).

(insert 1--2 paragraph summary of most important findings and how they
fit into the broader literature)

We expect that this article will be of interest to... (briefly describe target audience).

Thank you for considering this manuscript, and I hope you will find it suitable for publication in \textit{\journal}.

\end{newlfm}
\end{document}