### 2023-10-22 paper/main.tex

\documentclass[10pt]{article}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage[font=small,labelfont=bf]{caption}
\usepackage{geometry}
\usepackage[sort&compress, numbers]{natbib}
\usepackage{pxfonts}
\usepackage{graphicx}
\usepackage{setspace}
\usepackage{hyperref}
\usepackage{lineno}

\usepackage{xcolor}

% supplemental tables
\newcommand{\questions}{1}
\newcommand{\topics}{2}

% supplemental figures
\newcommand{\topicWordWeights}{1}
\newcommand{\topicWeights}{2}
\newcommand{\individualKnowledgeMapsA}{3}
\newcommand{\individualKnowledgeMapsB}{4}
\newcommand{\individualKnowledgeMapsC}{5}
\newcommand{\individualLearningMapsA}{6}
\newcommand{\individualLearningMapsB}{7}

% simple command for inline comments
\newcommand{\comment}[1]{}
% italicize section names in \nameref and place in \mbox to prevent bug when
% text is split across lines or pages
\NewCommandCopy{\oldnameref}{\nameref}
\renewcommand{\nameref}[1]{\mbox{\textit{\oldnameref{#1}}}}

\doublespacing
\linenumbers

\title{Text embedding models yield high-resolution insights into conceptual
knowledge from short multiple-choice quizzes}

\author{Paxton C. Fitzpatrick\textsuperscript{1},
Andrew C. Heusser\textsuperscript{1, 2}, and Jeremy R.
Manning\textsuperscript{1, *}\\\small{\textsuperscript{1}Department of Psychological and Brain Sciences}\\\small{Dartmouth College, Hanover, NH 03755, USA}\\\small{\textsuperscript{2}Akili Interactive Labs}\\\small{Boston, MA 02110, USA}\\\small{\textsuperscript{*}Corresponding author:
Jeremy.R.Manning@Dartmouth.edu}}

\date{}

\begin{document}
\maketitle

\begin{abstract}\noindent We develop a mathematical framework, based on natural
language processing models, for tracking and characterizing the acquisition of
conceptual knowledge. Our approach embeds each concept in a high-dimensional
representation space, where nearby coordinates reflect similar or related
concepts. We test our approach using behavioral data from participants who
answered small sets of multiple-choice quiz questions interleaved between
watching two course videos from the Khan Academy platform. We apply our
framework to the videos' transcripts and the text of the quiz questions to
quantify the content of each moment of video and each quiz question. We use
these embeddings, along with participants' quiz responses, to track how the
learners' knowledge changed after watching each video. Our findings show how a
small set of quiz questions may be used to obtain rich and meaningful
high-resolution insights into what each learner knows, and how their knowledge
changes over time as they learn.

\textbf{Keywords: education, learning, knowledge, concepts, natural language processing}

\end{abstract}

\section*{Introduction}

Suppose that a teacher had access to a complete, tangible ``map'' of everything
a student knows. Defining what such a map might even look like, let alone how it
might be constructed or filled in, is itself a non-trivial problem. But if a
teacher \textit{were} to gain access to such a map, how might it change their
ability to teach that student? Perhaps they might start by checking how well
the student knows the to-be-learned information already, or how much they know
about related concepts. For some students, they could potentially optimize
their teaching efforts to maximize efficiency by focusing primarily on
not-yet-known content. For other students (or other content areas), it might be
more effective to optimize for direct connections between already known content
and new material. Observing how the student's knowledge changed over time, in
response to their teaching, could also help to guide the teacher towards the
most effective strategy for that individual student.

A common approach to assessing a student's knowledge is to present them with a
set of quiz questions, calculate the proportion they answer correctly, and
provide them with feedback in the form of a simple numeric or letter grade.
While such a grade can provide \textit{some} indication of whether the student
has mastered the to-be-learned material, any univariate measure of performance
on a complex task sacrifices certain relevant information, risks conflating
underlying factors, and so on. For example, consider the relative utility of
the theoretical map described above that characterizes a student's knowledge in
detail, versus a single annotation saying that the student answered 85\% of
their quiz questions correctly, or that they received a `B'. Here, we show that
the same quiz data required to compute proportion-correct scores or letter
grades can instead be used to obtain far more detailed insights into what a
student knew at the time they took the quiz.

Designing and building procedures and tools for mapping out knowledge touches
on deep questions about what it means to learn. For example, how do we acquire
conceptual knowledge? Memorizing course lectures or textbook chapters by rote
can lead to the superficial \textit{appearance} of understanding the underlying
content, but achieving true conceptual understanding seems to require something
deeper and richer. Does conceptual understanding entail connecting newly
acquired information to the scaffolding of one's existing knowledge or
experience~\citep{BlayEtal06,CaraMaho03, ConsEtal16, DeacEtal04, SimoEtal04}?
Or weaving a lecture's atomic elements (e.g., its component words) into a
structured network that describes how those individual elements are
related~\citep{LeeChen22}? Conceptual understanding could also involve building
a mental model that transcends the meanings of those individual atomic elements
by reflecting the deeper meaning underlying the gestalt whole~\citep{Kint70,
Macl05, ScotEtal07}.

The difference between ``understanding'' and ``memorizing,'' as framed by
researchers in education, cognitive psychology, and cognitive neuroscience
(e.g.,~\citealp{Kato40, Gall00, ScotEtal07, HallGree08, Macl05}), has profound
analogs in the fields of natural language processing and natural language
understanding. For example, considering the raw contents of a document (e.g.,
its constituent symbols, letters, and words) might provide some clues as to
what the document is about, just as memorizing a passage might provide some
ability to answer simple questions about it. However, text embedding models
(e.g.,~\citealp{LandDuma97, DeerEtal90, BleiEtal03, BleiLaff06, MikoEtal13a,
CerEtal18, BrowEtal20}) also attempt to capture the deeper meaning
\textit{underlying} those atomic elements. These models consider not only the
co-occurrences of those elements within and across documents, but (in many
cases) also patterns in how those elements appear across different scales
(e.g., sentences, paragraphs, chapters, etc.), the temporal and grammatical
properties of the elements, and other high-level characteristics of how they
are used~\citep{Mann20, Mann21a}. According to these models, the deep
conceptual meaning of a document may be captured by a feature vector in a
high-dimensional representation space, wherein nearby vectors reflect
conceptually related documents. A model that succeeds at capturing an analogue
of ``understanding'' is able to assign nearby feature vectors to two
conceptually related documents, \textit{even when the specific words contained
in those documents have very little overlap}.

Given these insights, what form might a representation of the sum total of a
person's knowledge take (speculatively)? First, we might require a means of
systematically describing or representing the nearly infinite set of possible
things a person could know. Second, we might want to account for potential
associations between different concepts. For example, the concepts of ``fish''
and ``water'' might be associated in the sense that fish live in water. Third,
knowledge may have a critical dependency structure, such that knowing about a
particular concept might require first knowing about a set of other concepts.
For example, understanding the concept of a fish swimming in water first
requires understanding what fish and water \textit{are}. Fourth, as we learn,
our ``current state of knowledge'' should change accordingly. Learning new
concepts should both update our characterizations of ``what is known'' and also
unlock any now-satisfied dependencies of those newly learned concepts so that
they are ``tagged'' as available for future learning.

Here we develop a framework for modeling how conceptual knowledge is acquired
during learning. The central idea behind our framework is to use text embedding
models to define the coordinate systems of two maps: a \textit{knowledge
map} that describes the extent to which each concept is currently known, and
a \textit{learning map} that describes changes in knowledge over time. Each
location on these maps represents a single concept, and the maps' geometries
are defined such that related concepts are located nearby in space. We use this
framework to analyze and interpret behavioral data collected from an experiment
that had participants answer sets of multiple-choice questions about a series of
recorded course lectures.

Our primary research goal is to advance our understanding of what it means to
acquire deep, real-world conceptual knowledge. Traditional laboratory
approaches to studying learning and memory (e.g., list-learning studies) often
draw little distinction between memorization and understanding. Instead, these
studies typically focus on whether information is effectively encoded or
retrieved, rather than whether the information is \textit{understood}.
Approaches to studying conceptual learning, such as category learning
experiments, can begin to investigate the distinction between memorization and
understanding, often by training participants to distinguish arbitrary or
random features in otherwise meaningless categorized stimuli~\citep{ReilEtal82,
Este86a, Este86b, GlucEtal02, AshbMadd05, HulbNorm15}. However the objective of
real-world training, or learning from life experiences more generally, is often
to develop new knowledge that may be applied in \textit{useful} ways in the
future. In this sense, the gap between modern learning theories and modern
pedagogical approaches that inform classroom learning strategies is enormous:
most of our theories about \textit{how} people learn are inspired by
experimental paradigms and models that have only peripheral relevance to the
kinds of learning that students and teachers actually seek~\citep{Macl05,
HallGree08}. To help bridge this gap, our study uses course materials from real
online courses to inform, fit, and test models of real-world conceptual
learning. We also provide a demonstration of how our models can be used to
construct ``maps'' of what students know, and how their knowledge changes with
training. In addition to helping to visually capture knowledge (and changes in
knowledge), we hope that such maps might lead to real-world tools for improving
how we educate. Taken together, our work shows that existing course materials
and evaluative tools like short multiple-choice quizzes may be leveraged to
gain highly detailed insights into what students know and how they learn.

\section*{Results}

At its core, our main modeling approach is based around a simple assumption
that we sought to test empirically: all else being equal, knowledge about a
given concept is predictive of knowledge about similar or related concepts.
From a geometric perspective, this assumption implies that knowledge is
fundamentally ``smooth.'' In other words, as one moves through a space
representing an individual's knowledge (where similar concepts occupy nearby
coordinates), their ``level of knowledge'' should change relatively gradually.
To begin to test this smoothness assumption, we sought to track participants'
knowledge and how it changed over time in response to training. Two overarching
goals guide our approach. First, we want to gain detailed insights into what
learners know at different points in their training. For example, rather than
simply reporting on the proportions of questions participants answer correctly
(i.e., their overall performance), we seek estimates of their knowledge about a
variety of specific concepts. Second, we want our approach to be potentially
scalable to large numbers of diverse concepts, courses, and students. This
requires that the conceptual content of interest be discovered
\textit{automatically}, rather than relying on manually produced ratings or
labels.

\begin{figure}[tp]
    \centering
    \includegraphics[width=\textwidth]{figs/experiment}

    \caption{\textbf{Experimental paradigm.} Participants alternate between
    completing three 13-question multiple-choice quizzes and watching two Khan
    Academy lectures. Each quiz contains a mix of 5 questions about Lecture~1,
    5 questions about Lecture~2, and 3 questions about general physics
    knowledge. The specific questions reflected on each quiz, and the orders of
    each quiz's questions, were randomized across participants.}

    \label{fig:experiment}
\end{figure}

We asked participants in our study to complete brief multiple-choice quizzes
before, between, and after watching two lecture videos from the Khan
Academy~\citep{Khan04} platform (Fig.~\ref{fig:experiment}). The first lecture
video, entitled \textit{Four Fundamental Forces}, discussed the four
fundamental forces in physics: gravity, strong and weak interactions, and
electromagnetism. The second, entitled \textit{Birth of Stars}, provided an
overview of our current understanding of how stars form. We selected these
particular lectures to satisfy three general criteria. First, we wanted both
lectures to be accessible to a broad audience (i.e., with minimal prerequisite
knowledge) so as to limit the impact of prior training on participants'
abilities to learn from the lectures. To this end, we selected two introductory
videos that were intended to be viewed at the start of students' training in
their respective content areas. Second, we wanted the two lectures to have some
related content, so that we could test our approach's ability to distinguish
similar conceptual content. To this end, we chose two videos from the same Khan
Academy course domain, ``Cosmology and Astronomy.'' Third, we sought to
minimize dependencies and specific overlap between the videos. For example, we
did not want participants' abilities to understand one video to (directly)
influence their abilities to understand the other. To satisfy this last
criterion, we chose videos from two different lecture series (Lectures~1 and 2
were from the ``Scale of the Universe'' and ``Stars, Black Holes, and
Galaxies'' series, respectively).

We also wrote a set of multiple-choice quiz questions that we hoped would
enable us to evaluate participants' knowledge about each individual lecture,
along with related knowledge about physics concepts not specifically presented
in either video (see Supp. Tab.~\questions~for the full list of questions in
our stimulus pool). Participants answered questions randomly drawn from each
content area (Lecture~1, Lecture~2, and general physics knowledge) on each of
the three quizzes. Quiz~1 was intended to assess participants' ``baseline''
knowledge before training, Quiz~2 assessed knowledge after watching the
\textit{Four Fundamental Forces} video (i.e., Lecture~1), and Quiz~3 assessed
knowledge after watching the \textit{Birth of Stars} video (i.e., Lecture~2).

\begin{figure}[tp]
    \centering
    \includegraphics[width=\textwidth]{figs/sliding_windows}

    \caption{\textbf{Modeling course content.} \textbf{A. Building a document
    pool from sliding windows of text.} We decompose each lecture's transcript
    into a series of overlapping sliding windows. The full set of transcript
    snippets (across all windows) may be treated as a set of ``documents'' for
    training a text embedding model. \textbf{B. Constructing lecture content
    \textit{trajectories}.} After training the model on the sliding windows
    from both lectures, we transform each lecture into a ``trajectory'' through
    text embedding space by joining the embedding coordinates of successive
    sliding windows parsed from its transcript. \textbf{C. Embedding multiple
    lectures and questions in a shared space.} We apply the same model (trained
    on the two lectures' windows) to both lectures, along with the text of each
    question in our pool (Supp. Tab.~\questions), to project them into a shared
    text embedding space. This results in one trajectory per lecture and one
    coordinate for each question. Here, we have projected the 15-dimensional
    embeddings onto their first 3 principal components for visualization.}

    \label{fig:sliding-windows}
\end{figure}

To study in detail how participants' conceptual knowledge changed over the
course of the experiment, we first sought to model the conceptual content
presented to them at each moment throughout each of the two lectures. We
adapted an approach we developed in prior work~\citep{HeusEtal21} to identify
the latent themes in the lectures using a topic model~\citep{BleiEtal03}.
Briefly, topic models take as input a collection of text documents, and learn a
set of ``topics'' (i.e., latent themes) from their contents. Once fit, a topic
model can be used to transform arbitrary (potentially new) documents into sets
of ``topic proportions,'' describing the weighted blend of learned topics
reflected in their texts. We parsed automatically generated transcripts of the
two lectures into overlapping sliding windows, where each window contained the
text of the lecture transcript from a particular time span. We treated the set
of text snippets (across all of these windows) as documents to fit the model
(Fig.~\ref{fig:sliding-windows}A; see~\nameref{subsec:topic-modeling}).
Transforming the text from every sliding window with the model yielded a
number-of-windows by number-of-topics (15) topic-proportions matrix describing
the unique mixture of broad themes from both lectures reflected in each
window's text. Each window's ``topic vector'' (i.e., column of the
topic-proportions matrix) is analogous to a coordinate in a 15-dimensional
space whose axes are topics discovered by the model. Within this space, each
lecture's sequence of topic vectors (i.e., corresponding to its transcript's
overlapping text snippets across sliding windows) forms a \textit{trajectory}
that captures how its conceptual content unfolds over time
(Fig.~\ref{fig:sliding-windows}B). We resampled these trajectories to a
resolution of one topic vector for each second of video (i.e., 1~Hz).

We hypothesized that a topic model trained on transcripts of the two lectures
should also capture the conceptual knowledge probed by each quiz question. If
indeed the topic model could capture information about the deeper conceptual
content of the lectures (i.e., beyond surface-level details such as particular
word choices), then we should be able to recover a correspondence between each
lecture and questions \textit{about} each lecture. Importantly, such a
correspondence could not solely arise from superficial text matching between
lecture transcripts and questions, since the lectures and questions used
different words. Simply comparing the average topic weights from each lecture
and question set (averaging across time and questions, respectively) reveals a
striking correspondence (Supp. Fig.~\topicWeights). Specifically, the average
topic weights from Lecture~1 are strongly correlated with the average topic
weights from Lecture~1 questions ($r(13) = 0.809,~p < 0.001$, 95\% confidence
interval (CI)~$= [0.633,~0.962]$), and the average topic weights from Lecture~2
are strongly correlated with the average topic weights from Lecture~2 questions
($r(13) = 0.728,~p = 0.002$, 95\% CI~$= [0.456,~0.920]$). At the same time, the
average topic weights from the two lectures are \textit{negatively} correlated
with their non-matching question sets (Lecture~1 video vs.~Lecture~2 questions:
$r(13) = -0.547,~p = 0.035$, 95\% CI~$= [-0.812, ~-0.231]$; Lecture~2 video
vs.~Lecture~1 questions: $r(13) = -0.612,~p = 0.015$, 95\% CI~$=
[-0.874,~-0.281]$), indicating that the topic model also exhibits some degree
of specificity. The full set of pairwise comparisons between average topic
weights for the lectures and question sets is reported in Supplementary
Figure~\topicWeights.

\begin{figure}[tp]
    \centering
    \includegraphics[width=\textwidth]{figs/active-topics}

    \caption{\textbf{Lecture and question topic overlap.} \textbf{A. Topic
    weight variability}. The bar plots display the variance of each topic's
    weight across lecture timepoints (top row) and questions (bottom row);
    colors denote topics. The top-weighted words from the most ``expressive''
    (i.e., variable across observations) topic from each lecture are displayed
    in the upper right (orange: topic 2; yellow-green: topic 4). The
    top-weighted words from the full set of topics may be found in
    Supplementary Table~\topics. \textbf{B. Relationships between topic weight
    variability.} Pairwise correlations between the distributions of topic
    weight variance for each lecture and question set. Each row and column
    corresponds to a bar plot in Panel A.}

    \label{fig:topics}
\end{figure}

It is important to clarify that although we use topic model-derived embeddings
to \textit{characterize} the conceptual content of the lectures and questions,
we do not claim that the topic model itself \textit{understands} the conceptual
content of the lectures or questions. Rather, we view the topic model as a tool
for capturing the \textit{structure} of the conceptual content of the lectures
and questions in a way that enables us to capture, quantify, and track and
predict participants' knowledge.

Another, more sensitive, way of summarizing the conceptual content of the
lectures and questions is to look at \textit{variability} in how topics are
weighted over time and across different questions (Fig.~\ref{fig:topics}).
Intuitively, the variability in the expression of a given topic relates to how
much ``information''~\citep{Fish22} the lecture (or question set) reflects
about that topic. For example, suppose a given topic is weighted on heavily
throughout a lecture. That topic might be characteristic of some aspect or
property of the lecture \textit{overall} (conceptual or otherwise), but unless
the topic's weights changed in meaningful ways over time, the topic would be a
poor indicator of any \textit{specific} conceptual content in the lecture. We
therefore also compared the variances in topic weights (across time or
questions) between the lectures and questions. The variability in topic
expression (over time and across questions) was similar for the Lecture~1 video
and questions ($r(13) = 0.824,~p<0.001$, 95\% CI~$= [0.696,~0.973]$) and the
Lecture~2 video and questions ($r(13) = 0.801,~p<0.001$, 95\% CI~$=
[0.539,~0.958]$). Simultaneously, as reported in Figure~\ref{fig:topics}B, the
variability in topic expression across \textit{different} videos and
lecture-specific questions (i.e., Lecture~1 video vs.~Lecture~2 questions;
Lecture~2 video vs.~Lecture~1 questions) were negatively correlated, and
neither video's topic variability was reliably correlated with the topic
variability across general physics knowledge questions. Taken together, the
analyses reported in Figure~\ref{fig:topics} and Supplementary
Figure~\topicWeights~indicate that a topic model fit to the videos' transcripts
can also reveal correspondences (at a coarse scale) between the lectures and
questions.

While an individual lecture may be organized around a single broad theme at a
coarse scale, at a finer scale, each moment of a lecture typically covers a
narrower range of content. Given the correspondence we found between the
variability in topic expression across moments of each lecture and questions
from its corresponding set (Fig.~\ref{fig:topics}), we wondered whether the
text embedding model might additionally capture these conceptual relationships
at a finer scale. For example, if a particular question asks about the content
from one small part of a lecture, we wondered whether the text embeddings could
be used to automatically identify the ``matching'' moment(s) in the lecture. To
explore this, we computed the correlation between each question's topic weights
and the topic weights for each second of its corresponding lecture, and found
that each question appeared to be temporally specific
(Fig.~\ref{fig:question-correlations}). In particular, most questions' topic
vectors were maximally correlated with a well-defined (and relatively narrow)
range of timepoints from their corresponding lectures, and the correlations
fell off sharply outside of that range. We also qualitatively examined the
best-matching intervals for each question by comparing the question's text to
the text of the most-correlated parts of the lectures. Despite that the
questions were excluded from the text embedding model's training set, in
general we found (through manual inspection) a close correspondence between the
conceptual content that each question probed and the content covered by the
best-matching moments of the lectures. Two representative examples are shown at
the bottom of Figure~\ref{fig:question-correlations}.

\begin{figure}[t]
    \centering
    \includegraphics[width=\textwidth]{figs/lecture-question-similarity}

    \caption{\textbf{Which parts of each lecture are captured by each
    question?} Each panel displays timeseries plots showing how each question's
    topic vector correlates with each video timepoint's topic vector (Panel
    \textbf{A.}: correlations for the \textit{Four Fundamental Forces} lecture
    and associated questions; Panel \textbf{B.}: correlations for the
    \textit{Birth of Stars} lecture and associated questions). The colors
    denote question identities. The diamonds in each panel denote the moment of
    peak correlation between the indicated question and the lecture trajectory.
    The associated questions' text and snippets of the lectures' transcripts
    from the surrounding 30 seconds, are displayed at the bottom of the
    figure.}

    \label{fig:question-correlations}
\end{figure}

The ability to quantify how much each question is ``asking about'' the content
from each moment of the lectures could enable high-resolution insights into
participants' knowledge. Traditional approaches to estimating how much a
student ``knows'' about the content of a given lecture entail computing the
proportion of correctly answered questions. But if two students receive
identical scores on an exam, might our modeling framework help us to gain more
nuanced insights into the \textit{specific} content that each student has
mastered (or failed to master)? For example, a student who misses three
questions that were all about the same concept (e.g., concept $A$) will have
gotten the same \textit{proportion} of questions correct as another student who
missed three questions about three \textit{different} concepts (e.g., $A$, $B$,
and $C$). But if we wanted to help these two students fill in the ``gaps'' in
their understandings, we might do well to focus specifically on concept $A$ for
the first student, but to also add in materials pertaining to concepts $B$ and
$C$ for the second student. In other words, raw ``proportion-correct'' measures
may capture \textit{how much} a student knows, but not \textit{what} they know.
We wondered whether our modeling framework might enable us to (formally and
automatically) infer participants' knowledge at the scale of individual
concepts (e.g., as captured by a single moment of a lecture).

We developed a simple formula (Eqn.~\ref{eqn:prop}) for using a participant's
responses to a small set of multiple-choice questions to estimate how much the
participant ``knows'' about the concept reflected by any arbitrary coordinate,
$x$, in text embedding space (e.g., the content reflected by any moment in a
lecture they had watched; see \nameref{subsec:traces}). Essentially, the
estimated knowledge at coordinate $x$ is given by the weighted average
proportion of quiz questions the participant answered correctly, where the
weights reflect how much each question is ``about'' the content at $x$. When we
apply this approach to estimate the participant's knowledge about the content
presented in each moment of each lecture, we can obtain a detailed timecourse
describing how much ``knowledge'' the participant has about any part of the
lecture. As shown in Figure~\ref{fig:knowledge-timeseries}A and C, we can apply
this approach separately for the questions from each quiz participants took
throughout the experiment. From just a few questions per quiz
(see~\nameref{subsec:traces}), we obtain a high-resolution snapshot (at the
time each quiz was taken) of what the participants knew about any moment's
content, from either of the two lectures they watched (comprising a total of
1,100 samples across the two lectures).

\begin{figure}[tp]
    \centering
    \includegraphics[width=0.7\textwidth]{figs/content-mastery}

    \caption{\textbf{Estimating moment-by-moment knowledge acquisition.}
    \textbf{A. Moment-by-moment knowledge about the \textit{Four Fundamental
    Forces}}. Each trace displays the weighted proportion of correctly answered
    questions about the content reflected in each moment of the lecture (see
    \nameref{subsec:traces}), using responses from a single quiz (color). The
    traces are averaged across participants. \textbf{B. Average estimated
    knowledge about the \textbf{Four Fundamental Forces}.} Each bar displays
    the across-timepoint average knowledge, estimated using the responses to
    one quiz's questions. \textbf{C. Moment-by-moment knowledge about the
    \textit{Birth of Stars}}. The panel is in the same format as Panel A, but
    here the knowledge estimates are for the moment-by-moment content of the
    \textit{Birth of Stars} lecture. \textbf{D. Average estimated knowledge
    about the \textit{Birth of Stars}.} The panel is in the same format as
    Panel B, but here the knowledge estimates are for the content of the
    \textit{Birth of Stars} lecture. All panels: error ribbons and error bars
    denote 95\% confidence intervals, estimated across participants.}

    \label{fig:knowledge-timeseries}
\end{figure}

While the timecourses in Figure~\ref{fig:knowledge-timeseries}A and C provide
detailed \textit{estimates} about participants' knowlege, these estimates are
of course only \textit{useful} to the extent that they accurately reflect what
participants actually know. As one sanity check, we anticipated that the
knowledge estimates should reflect a content-specific ``boost'' in
participants' knowledge after watching each lecture. In other words, if
participants learn about each lecture's content when they watch each lecture,
the knowledge estimates should capture that. After watching the \textit{Four
Fundamental Forces} lecture, participants should exhibit more knowledge for the
content of that lecture than they had before, and that knowledge should persist
for the remainder of the experiment. Specifically, knowledge about that
lecture's content should be relatively low when estimated using Quiz~1
responses, but should increase when estimated using Quiz~2 or 3 responses
(Fig.~\ref{fig:knowledge-timeseries}B). Indeed, we found that participants'
estimated knowledge about the content of the \textit{Four Fundamental Forces}
was substantially higher on Quiz~2 versus Quiz~1 ($t(49) = 8.764,~p < 0.001$)
and on Quiz~3 versus Quiz~1 ($t(49) = 10.519,~p < 0.001$). We found no reliable
differences in estimated knowledge about that lecture's content on Quiz~2
versus 3 ($t(49) = 0.160,~p = 0.874$). Similarly, we hypothesized (and
subsequently confirmed) that participants should show greater estimated
knowledge about the content of the \textit{Birth of Stars} lecture after
(versus before) watching it (Fig.~\ref{fig:knowledge-timeseries}D).
Specifically, since participants watched that lecture after taking Quiz~2 (but
before Quiz~3), we hypothesized that their knowledge estimates should be
relatively low on Quizzes~1 and 2, but should show a ``boost'' on Quiz~3.
Consistent with this prediction, we found no reliable differences in estimated
knowledge about the \textit{Birth of Stars} lecture content on Quizzes~1 versus
2 ($t(49) = 1.013,~p = 0.316$), but the estimated knowledge was substantially
higher on Quiz~3 versus 2 ($t(49) = 10.561,~p < 0.001$) and Quiz~3 versus 1
($t(49) = 8.969,~p < 0.001$).

If we are able to accurately estimate a participant's knowledge about the
content tested by a given question, our estimates of their knowledge should
carry some predictive information about whether the participant is likely to
answer that question correctly or incorrectly. We developed a statistical
approach to test this claim. For each question, in turn, we used
Equation~\ref{eqn:prop} to estimate each participant's knowledge at the given
question's embedding space coordinate, using all \textit{other} questions that
participant answered on the same quiz. For each quiz, we grouped these
estimates into two distributions: one for the estimated knowledge at the
coordinates of \textit{correctly} answered questions, and another for the
estimated knowledge at the coordinates of \textit{incorrectly} answered
questions (Fig.~\ref{fig:predictions}). We then used independent samples
$t$-tests to compare the means of these distributions of estimated knowledge.

\begin{figure}[tp]
    \centering
    \includegraphics[width=0.7\textwidth]{figs/predict-knowledge-questions}

    \caption{\textbf{Estimating knowledge at the embedding coordinates of
    held-out questions.} Separately for each quiz (panel), we plot the
    distributions of predicted knowledge at the embedding coordinates of each
    held-out correctly (blue) or incorrectly (red) answered question. The
    $t$-tests reported in each panel are between the distributions of estimated
    knowledge at the coordinates of correctly versus incorrectly answered
    held-out questions.}

    \label{fig:predictions}
\end{figure}

For the initial quizzes participants took (prior to watching either lecture),
participants' estimated knowledge tended to be low overall, and relatively
unstructured (Fig.~\ref{fig:predictions}, left panel). When we held out
individual questions and estimated their knowledge at the held-out questions'
embedding coordinates, we found no reliable differences in the estimates when
the held-out question had been correctly versus incorrectly answered ($t(633) =
0.577,~p = 0.564$). This reflects a floor effect: when knowledge is low
everywhere, there is little signal to differentiate between what is known
versus unknown. After watching the first lecture, estimated knowledge for
held-out correctly answered questions (from the second quiz;
Fig.~\ref{fig:predictions}, middle panel) exhibited a positive shift relative
to held-out incorrectly answered questions ($t(633) = 3.961,~p < 0.001$). This
second quiz provides the maximally sensitive test for our knowledge
predictions, since (if knowledge is estimated accurately) participants' Quiz~2
responses should demonstrate specific knowledge about Lecture~1 content, but
knowledge about Lecture~2 and general physics concepts should be roughly
unchanged from before they watched Lecture~1. After watching the second
lecture, estimated knowledge (from the third quiz; Fig.~\ref{fig:predictions},
right panel) for \textit{all} questions exhibited a positive shift. However,
the estimated knowledge for held-out correctly answered questions remained
greater than that for held-out incorrectly answered questions ($t(628) =
2.045,~p = 0.041$). This third contrast reflects a ceiling effect: when
knowledge is relatively high everywhere, the signal differentiating what is
known versus unknown is relatively weak. Taken together, this set of analyses
demonstrates that our knowledge prediction framework is most informative when
participants exhibit variability in their knowledge of the content captured by
the text embedding model.

\begin{figure}[t]
    \centering
    \includegraphics[width=\textwidth]{figs/knowledge-smoothness}

    \caption{\textbf{Caption title.} Caption content.}

    \label{fig:smoothness}
\end{figure}

Knowledge estimates need not be limited to the content of the lectures. As
illustrated in Figure~\ref{fig:knowledge-maps}, our general approach to
estimating knowledge from a small number of quiz questions may be extended to
\textit{any} content, given its text embedding coordinate. To visualize how
knowledge ``spreads'' through text embedding space to content beyond the
lectures participants watched, we first fit a new topic model to the lectures'
sliding windows with (up to)~$k = 100$~topics. Conceptually, increasing the
number of topics used by the model functions to increase the ``resolution'' of
the embedding space, providing a greater ability to estimate knowledge for
content that is highly similar to (but not precisely the same as) that
contained in the two lectures. Aside from increasing the number of topics from
15 to 100, all other procedures and model parameters were carried over from the
preceding analyses. As in our other analyses, we resampled each lecture's topic
trajectory to 1~Hz and projected each question into a shared text embedding
space.

\begin{figure}[tp]
    \centering
    \includegraphics[width=\textwidth]{figs/knowledge_and_learning_maps}

    \caption{\textbf{Mapping out the geometry of knowledge and learning.}
    \textbf{A. Average ``knowledge maps'' estimated using each quiz.} Each map
    displays a 2D projection of the estimated knowledge about the content
    reflected by \textit{all} regions of topic space (see
    \nameref{subsec:knowledge-maps}). The topic trajectories of the two
    lectures are indicated by dotted lines (blue: Lecture 1; green: Lecture 2),
    and the coordinates of each question are indicated by dots (light blue:
    Lecture 1-related; light green: Lecture 2-related; purple: general physics
    knowledge). Each map reflects an average across all participants. For
    individual participants' maps, see Supplementary
    Figures~\individualKnowledgeMapsA,~\individualKnowledgeMapsB,
    and~\individualKnowledgeMapsC. \textbf{B. Average ``learning maps''
    estimated between each successive pair of quizzes.} The learning maps
    follow the same general format as the knowledge maps in Panel A, but here
    the shading at each coordinate indicates the \textit{difference} between
    the corresponding coordinates in the indicated \textit{pair} of knowledge
    maps---i.e., how much the estimated knowledge ``changed'' between the two
    quizzes. Each map reflects an average across all participants. For
    individual participants' maps, see Supplementary
    Figures~\individualLearningMapsA~and~\individualLearningMapsB. \textbf{C.
    Word clouds for sampled points in topic space.} Each word cloud displays
    the weighted blend of words underlying the topic proportions represented at
    the corresponding colored star's location on the maps. In each word cloud,
    the words' relative sizes correspond to their relative weights at the
    starred location, and their colors indicate their relative weights in the
    \textit{Four Fundamental Forces} (blue) versus \textit{Birth of Stars}
    (green) lectures, on average, across all timepoints' topic vectors.}

    \label{fig:knowledge-maps}
    \end{figure}

We projected the resulting 100-dimensional topic vectors (for each second of
video and each quiz question) onto a shared 2-dimensional plane (see
\nameref{subsec:knowledge-maps}). Next, we sampled points from a $100 \times
100$ grid of coordinates that evenly tiled a rectangle enclosing the 2D
projections of the videos and questions. We used
Equation~\ref{eqn:rbf-knowledge} to estimate participants' knowledge at each of
these 10,000 sampled locations, and averaged these estimates across
participants to obtain an estimated average \textit{knowledge map}
(Fig.~\ref{fig:knowledge-maps}A). Intuitively, the knowledge map constructed
from a given quiz's responses provides a visualization of how ``much''
participants knew about any content expressible by the fitted text embedding
model at the point in time when they completed that quiz.

Several features of the resulting knowledge maps are worth noting. The average
knowledge map estimated from Quiz~1 responses (Fig.~\ref{fig:knowledge-maps}A,
leftmost map) shows that participants tended to have relatively little
knowledge about any parts of the text embedding space (i.e., the shading is
relatively dark everywhere). The knowledge map estimated from Quiz~2 responses
shows a marked increase in knowledge on the left side of the map (around
roughly the same range of coordinates traversed by the \textit{Four Fundamental
Forces} lecture, indicated by the dotted blue line). In other words,
participants' estimated increase in knowledge is localized to conceptual
content that is nearby (i.e., related to) the content from the lecture they
watched prior to taking Quiz~2. This localization is non-trivial: these
knowledge estimates are informed only by the embedded coordinates of the
\textit{quiz questions}, not by the embeddings of either lecture (see
Eqn.~\ref{eqn:rbf-knowledge}). Finally, the knowledge map estimated from Quiz 3
responses shows a second increase in knowledge, localized to the region
surrounding the embedding of the \textit{Birth of Stars} lecture participants
watched immediately prior to taking Quiz~3.

Another way of visualizing these content-specific increases in knowledge after
participants viewed each lecture is displayed in
Figure~\ref{fig:knowledge-maps}B. Taking the point-by-point difference between
the knowledge maps estimated from responses to a successive pair of quizzes
yields a \textit{learning map} that describes the \textit{change} in knowledge
estimates from one quiz to the next. These learning maps highlight that the
estimated knowledge increases we observed across maps were specific to the
regions around the embeddings of each lecture, in turn.

Because the 2D projection we used to construct the knowledge and learning maps
is invertible, we may gain additional insights into these maps' meaning by
reconstructing the original high-dimensional topic vector for any location on
the map we are interested in. For example, this could serve as a useful tool
for an instructor looking to better understand which content areas a student
(or a group of students) knows well (or poorly). As a demonstration, we show
the top-weighted words from the blends of topics reconstructed from three
example locations on the maps (Fig.~\ref{fig:knowledge-maps}C): one point near
the \textit{Four Fundamental Forces} embedding (yellow), a second point near
the \textit{Birth of Stars} embedding (orange), and a third point between the
two lectures' embeddings (pink). As shown in the word clouds in the panel, the
top-weighted words at the example coordinate near the \textit{Four Fundamental
Forces} embedding tended to be weighted more heavily by the topics expressed in
that lecture. Similarly, the top-weighted words at the example coordinate near
the \textit{Birth of Stars} embedding tended to be weighted more heavily by the
topics expressed in \textit{that} lecture. And the top-weighted words at the
example coordinate between the two lectures' embeddings show a roughly even mix
of words most strongly associated with each lecture.

\section*{Discussion}

We developed a computational framework that uses short multiple-choice quizzes
to gain nuanced insights into what learners know and how their knowledge
changes with training. First, we show that our approach can automatically match
the conceptual knowledge probed by individual quiz questions to the
corresponding moments in lecture videos when those concepts were presented
(Fig.~\ref{fig:question-correlations}). Next, we demonstrate how we can
estimate moment-by-moment ``knowledge traces'' that reflect the degree of
knowledge participants have about each video's time-varying content, and
capture temporally specific increases in knowledge after viewing each lecture
(Fig.~\ref{fig:knowledge-timeseries}). We also show that these knowledge
estimates can generalize to held-out questions (Fig.~\ref{fig:predictions}).
Finally, we use our framework to construct visual maps that provide snapshot
estimates of how much participants know about any concept within the scope of
our text embedding model, and how much their knowledge of those concepts
changes with training (Fig.~\ref{fig:knowledge-maps}).

We view our work as making several contributions to the study of how people
acquire conceptual knowledge. First, from a methodological standpoint, our
modeling framework provides a systematic means of mapping out and
characterizing knowledge in maps that have infinite (arbitrarily many) numbers
of coordinates, and of ``filling out'' those maps using relatively small
numbers of multiple choice quiz questions. Our experimental finding that we can
use these maps to predict responses to held-out questions
(Fig.~\ref{fig:predictions}) also has important psychological implications. One
such psychological implication is that concepts that are assigned to nearby
coordinates by the text embedding model also appear to be ``known to a similar
extent'' (as reflected by participants' responses to held-out questions). This
suggests that participants also \textit{conceptualize} similarly the content
reflected by nearby embedding coordinates. A second psychological implication
is that something about how participants' knowledge ``spreads'' across concepts
is being captured by the knowledge maps we are inferring from their quiz
responses (e.g., Fig.~\ref{fig:smoothness}). In other words, our study shows
that knowledge about a given concept implies knowledge about related concepts.

Over the past several years, the global pandemic has forced many educators to
suddenly adapt to teaching remotely~\citep{MoseEtal21, ShimLee20, KawaEtal21,
Whal20}. This change in world circumstances is happening alongside (and perhaps
accelerating) geometric growth in the availability of high-quality online
courses from platforms such as Khan Academy~\citep{Khan04},
Coursera~\citep{Youn12}, EdX~\citep{Kolo13}, and others~\citep{RhoaEtal13}.
Continued expansion of the global internet backbone and improvements in
computing hardware have also facilitated improvements in video streaming,
enabling videos to be easily shared and viewed by increasingly large segments
of the world's population. This exciting time for online course instruction
provides an opportunity to re-evaluate how we, as a global community, educate
ourselves and each other. For example, we can ask: what defines an effective
course or training program? Which aspects of teaching might be optimized and/or
augmented by automated tools? How and why do learning needs and goals vary
across people? How might we lower barriers of access to a high-quality
education?

Alongside these questions, there is a growing desire to extend existing
theories beyond the domain of lab testing rooms and into real
classrooms~\citep{Kauf03}. In part, this has led to a recent resurgence of
``naturalistic'' or ``observational'' experimental paradigms that attempt to
better reflect more ethologically valid phenomena that are more directly
relevant to real-world situations and behaviors~\citep{NastEtal20}. In turn,
this has brought new challenges in data analysis and interpretation. A key step
towards solving these challenges will be to build explicit models of real-world
scenarios and how people behave in them (e.g., models of how people learn
conceptual content from real-world courses, as in our current study). A second
key step will be to understand which sorts of signals derived from behaviors
and/or other measurements (e.g., neurophysiological data;~\citealp{NguyEtal22,
MeshEtal20, PoulEtal17, BeviEtal19, DikkEtal17}) might help to inform these
models. A third major step will be to develop and employ reliable ways of
evaluating the complex models and data that are a hallmark of naturalistic
paradigms.

Beyond specifically predicting what people \textit{know}, the fundamental ideas
we develop here also relate to the notion of ``theory of mind'' of other
individuals~\citep{GoldWinn12, KansEtal15, Melt11}. Considering others' unique
perspectives, prior experiences, knowledge, goals, etc., can help us to more
effectively interact and communicate~\citep{ShaoEtal18, StepBaer06, Ratk18}.
One could imagine future extensions of our work (e.g., analogous to the
knowledge and learning maps shown in Fig.~\ref{fig:knowledge-maps}), that
attempt to characterize how well-aligned different people's knowledge bases or
backgrounds are. In turn, this might be used to model how knowledge (or other
forms of communicable information) flows not just between teachers and
students, but between friends having a conversation, individuals on a first
date, participants at a business meeting, doctors and patients, experts and
non-experts, political allies or adversaries, and more. For example, the extent
to which two people's knowledge maps ``match'' or ``align'' in a given region
of text embedding space might serve as a predictor of how effectively they will
be able to communicate about the corresponding conceptual content.

Ultimately, our work suggests a rich new line of questions about the geometric
``form'' of knowledge, how knowledge changes over time, and how we might map
out the full space of what an individual knows. Our finding that detailed
estimates about knowledge may be obtained from short quizzes shows one way that
traditional approaches to evaluation in education may be extended. We hope that
these advances might help pave the way for new approaches to teaching or
delivering educational content that are tailored to individual students'
learning needs and goals.

\section*{Materials and methods}

\subsection*{Participants}

We enrolled a total of 50 Dartmouth undergraduate students in our study.
Participants received optional course credit for enrolling. We asked each
participant to complete a demographic survey that included questions about
their age, gender, native spoken language, ethnicity, race, hearing, color
vision, sleep, coffee consumption, level of alertness, and several aspects of
their educational background and prior coursework.

Participants' ages ranged from 18 to 22 years (mean: 19.52 years; standard
deviation: 1.09 years). A total of 15 participants reported their gender as
male and 35 participants reported their gender as female. A total of 49
participants reported their native language as ``English'' and 1 reported
having another native language. A total of 47 participants reported their
ethnicity as ``Not Hispanic or Latino'' and three reported their ethnicity as
``Hispanic or Latino.'' Participants reported their races as White (32
participants), Asian (14 participants), Black or African American (5
participants), American Indian or Alaska Native (1 participant), and Native
Hawaiian or Other Pacific Islander (1 participant). (Note that some
participants selected multiple racial categories.)

A total of 49 participants reporting having normal hearing and 1 participant
reported having some hearing impairment. A total of 49 participants reported
having normal color vision and 1 participant reported being color blind.
Participants reported having had, on the night prior to testing, 2--4 hours of
sleep (1 participant), 4--6 hours of sleep (9 participants), 6--8 hours of
sleep (35 participants), or $8+$ hours of sleep (5 participants). They reported
having consumed, on the same day and leading up to their testing session, 0
cups of coffee (38 participants), 1 cup of coffee (10 participants), 3 cups of
coffee (1 participant), or $4+$ cups of coffee (1 participant).

No participants reported that their focus was currently impaired (e.g., by
drugs or alcohol). Participants reported their current level of alertness, and
we converted their responses to numerical scores as follows: ``very sluggish''
(-2), ``a little sluggish'' (-1), ``neutral'' (0), ``fairly alert'' (1), and
``very alert'' (2). Across all participants, a range of alertness levels were
reported (range: -2--1; mean: -0.10; standard deviation: 0.84).

Participants reported their undergraduate major(s) as ``social sciences'' (28
participants), ``natural sciences'' (16 participants), ``professional'' (e.g.,
pre-med or pre-law; 8 participants), ``mathematics and engineering'' (7
participants), ``humanities'' (4 participants), or ``undecided'' (3
participants). Note that some participants selected multiple categories for
their undergraduate major(s). We also asked participants about the courses they
had taken. In total, 45 participants reported having taken at least one Khan
Academy course in the past, and 5 reported not having taken any Khan Academy
courses. Of those who reported having watched at least one Khan Academy course,
7 participants reported having watched 1--2 courses, 11 reported having watched
3--5 courses, 8 reported having watched 5--10 courses, and 19 reported having
watched 10 or more courses. We also asked participants about the specific
courses they had watched, categorized under different subject areas. In the
``Mathematics'' area, participants reported having watched videos on AP
Calculus AB (21 participants), Precalculus (17 participants), Algebra 2 (14
participants), AP Calculus BC (12 participants), Trigonometry (11
participants), Algebra 1 (10 participants), Geometry (8 participants),
Pre-algebra (7 participants), Multivariable Calculus (5 participants),
Differential Equations (5 participants), Statistics and Probability (4
participants), AP Statistics (2 participants), Linear Algebra (2 participants),
Early Math (1 participant), Arithmetic (1 participant), and other videos not
listed in our survey (5 participants). In the ``Science and engineering'' area,
participants reported having watched videos on Chemistry, AP Chemistry, or
Organic Chemistry (21 participants); Physics, AP Physics I, or AP Physics II
(18 participants); Biology, AP Biology; or High school Biology (15
participants); Health and Medicine (1 participant); or other videos not listed
in our survey (5 participants). We also asked participants whether they had
specifically seen the videos used in our experiment. Of the 45 participants who
reported having having taken at least one Khan Academy course in the past, 44
participants reported that they had not watched the \textit{Four Fundamental
Forces} video, and 1 participant reported that they were not sure whether they
had watched it. All participants reported that they had not watched the
\textit{Birth of Stars} video. When we asked participants about non-Khan
Academy online courses, they reported having watched or taken courses on
Mathematics (15 participants), Science and engineering (11 participants), Test
preparation (9 participants), Economics and finance (3 participants), Arts and
humanities (2 participants), Computing (2 participants), and other categories
not listed in our survey (17 participants). Finally, we asked participants
about in-person courses they had taken in different subject areas. They
reported taking courses in Mathematics (38 participants), Science and
engineering (37 participants), Arts and humanities (34 participants), Test
preparation (27 participants), Economics and finance (26 participants),
Computing (14 participants), College and careers (7 participants), or other
courses not listed in our survey (6 participants).

\subsection*{Experiment}\label{subsec:experiment}

We hand-selected two course videos from the Khan Academy platform: \textit{Four
Fundamental Forces} (an introduction to gravity, electromagnetism, the weak
nuclear force, and the strong nuclear force; duration: 10 minutes and 29
seconds) and \textit{Birth of Stars} (an introduction to how stars are formed;
duration: 7 minutes and 57 seconds). We then hand-created 39 multiple-choice
questions: 15 about the conceptual content of \textit{Four Fundamental Forces}
(i.e., Lecture~1), 15 about the conceptual content of \textit{Birth of Stars}
(i.e., Lecture~2), and 9 questions that tested for general conceptual knowledge
about basic physics (covering material that was not presented in either video).
We note that estimating the specific ``amount'' of conceptual understanding
that each question ``requires'' to answer is somewhat subjective, and might
even come down to the ``strategy'' a given participant uses to answer the
question at that particular moment. The full set of questions and answer
choices may be found in Supplementary Table~\questions.

Over the course of the experiment, participants completed three 13-question
multiple-choice quizzes: the first before viewing Lecture~1, the second between
Lectures~1 and 2, and the third after viewing Lecture~2 (see
Fig.~\ref{fig:experiment}). The questions appearing on each quiz, for each
participant, were randomly chosen from the full set of 39, with the constraints
that (a) each quiz contain exactly 5 questions about Lecture~1, 5 questions
about Lecture~2, and 3 questions about general physics knowledge, and (b) each
question appear exactly once for each participant. The orders of questions on
each quiz, and the orders of answer options for each question, were also
randomized. Our experimental protocol was approved by the Committee for the
Protection of Human Subjects at Dartmouth College. We used this experiment to
develop and test our computational framework for estimating knowledge and
learning.

\subsection*{Analysis}

\subsubsection*{Constructing text embeddings of multiple lectures and questions}\label{subsec:topic-modeling}

We adapted an approach we developed in prior work~\citep{HeusEtal21} to embed
each moment of the two lectures and each question in our pool in a common
representational space. Briefly, our approach uses a topic model (Latent
Dirichlet Allocation;~\citealp{BleiEtal03}), trained on a set of documents, to
discover a set of (up to) $k$ ``topics'' or ``themes.'' Formally, each topic is
defined as a distribution of weights over each word in the model's vocabulary
(i.e., the union of all unique words, across all documents, excluding ``stop
words.''). Conceptually, each topic is intended to give larger weights to words
that are semantically related or tend to co-occur in the same documents. After
fitting a topic model, each document in the training set, or any \textit{new}
document that contains at least some of the words in the model's vocabulary,
may be represented as a $k$-dimensional vector describing how much the document
(most probably) reflects each topic. To select an appropriate $k$ for our
model, we identified the minimum number of topics that yielded at least one
``unused'' topic (i.e., in which all words in the vocabulary were assigned
uniform weights) after training. This indicated that the number of topics was
sufficient to capture the set of latent themes present in the two lectures
(from which we constructed our document corpus, as described below). We found
this value to be $k = 15$ topics. The distribution of weights over words in the
vocabulary for each discovered topic is shown in Supplementary
Figure~\topicWordWeights, and each topic's top-weighted words may be found in
Supplementary Table~\topics.

As illustrated in Figure~\ref{fig:sliding-windows}A, we start by building up a
corpus of documents using overlapping sliding windows that span each video's
transcript. Khan Academy provides professionally created, manual transcriptions
of all videos for closed captioning. However, such transcripts would not be
readily available in all contexts to which our framework could potentially be
applied. Khan Academy videos are hosted on the YouTube platform, which
additionally provides automated captions. We opted to use these automated
transcripts (which, in prior work, we have found to be of sufficiently
near-human quality to yield reliable data in behavioral
studies;~\citealp{ZimaEtal18}) when developing our framework in order to make
it more directly extensible and adaptable by others in the future.

We fetched these automated transcripts using the
\texttt{youtube-transcript-api} Python package~\citep{Depo18}. The transcripts
consisted of one timestamped line of text for every few seconds (mean: 2.34~s;
standard deviation: 0.83~s) of spoken content in the video (i.e., corresponding
to each individual caption that would appear on-screen if viewing the lecture
via YouTube, and when those lines would appear). We defined a sliding window
length of (up to) $w = 30$ transcript lines, and assigned each window a
timestamp corresponding to the midpoint between the timestamps for its first
and last lines. These sliding windows ramped up and down in length at the
beginning and end of each transcript, respectively. In other words, each
transcript's first sliding window covered only its first line, the second
sliding window covered the first two lines, and so on. This ensured that each
line from the transcripts appeared in the same number ($w$) of sliding windows.
After performing various standard text preprocessing (e.g., normalizing case,
lemmatizing, removing punctuation and stop-words), we treated the text from
each sliding window as a single ``document,'' and combined these documents
across the two videos' windows to create a single training corpus for the topic
model.

After fitting a topic model to the two videos' transcripts, we could use the
trained model to transform arbitrary (potentially new) documents into
$k$-dimensional topic vectors. A convenient property of these topic vectors is
that documents that reflect similar blends of topics (i.e., documents that
reflect similar themes, according to the model) will yield similar coordinates
(in terms of correlation, cosine similarity, Kullback-Leibler divergence,
Euclidean distance, or other geometric measures). In general, the similarity
between different documents' topic vectors may be used to characterize the
similarity in conceptual content between the documents.

We transformed each sliding window's text into a topic vector, and then used
linear interpolation (independently for each topic dimension) to resample the
resulting timeseries to one vector per second. We also used the fitted model to
obtain topic vectors for each question in our pool (see Supp. Tab.~\questions).
Taken together, we obtained a \textit{trajectory} for each video, describing
its path through topic space, and a single coordinate for each question
(Fig.~\ref{fig:sliding-windows}C). Embedding both videos and all of the
questions using a common model enables us to compare the content from different
moments of videos, compare the content across videos, and estimate potential
associations between specific questions and specific moments of video.

\subsubsection*{Estimating dynamic knowledge traces}\label{subsec:traces}

We used the following equation to estimate each participant's knowledge about
timepoint $t$ of a given lecture, $\hat{k}(t)$:
\begin{equation}
    \hat{k}\left(f(t, L)\right) = \frac{\sum_{i \in \mathrm{correct}}\mathrm{ncorr}\left(f(t, L), f(i, Q)\right)}{\sum_{j = 1}^N \mathrm{ncorr}\left(f(t, L), f(j, Q)\right)},
    \label{eqn:prop}
\end{equation}
where

\begin{equation}
    \mathrm{ncorr}(x, y) = \frac{\mathrm{corr}(x, y) - \mathrm{mincorr}}{\mathrm{maxcorr} - \mathrm{mincorr}},
\end{equation}
and where $\mathrm{mincorr}$ and $\mathrm{maxcorr}$ are the minimum and maximum
correlations between any lecture timepoint and question, taken over all
timepoints in the given lecture, and all five questions \textit{about} that lecture appearing on the given quiz.
We also define $f(s, \Omega)$ as the
$s$\textsuperscript{th} topic vector from the set of topic vectors $\Omega$.
Here $t$ indexes the set of lecture topic vectors, $L$, and $i$ and $j$ index
the topic vectors of questions used to estimate the knowledge trace, $Q$. Note
that ``correct'' denotes the set of indices of the questions the participant
answered correctly on the given quiz.

Intuitively, $\mathrm{ncorr}(x, y)$ is the correlation between two topic
vectors (e.g., the topic vector from one timepoint in a lecture, $x$, and the
topic vector for one question, $y$), normalized by the minimum and maximum
correlations (across all timepoints $t$ and questions $Q$) to range between 0
and 1, inclusive. Equation~\ref{eqn:prop} then computes the weighted average
proportion of correctly answered questions about the content presented at
timepoint $t$, where the weights are given by the normalized correlations
between timepoint $t$'s topic vector and the topic vectors for each question.
The normalization step (i.e., using $\mathrm{ncorr}$ instead of the raw
correlations) insures that every question contributes some non-negative amount
to the knowledge estimate.

\subsubsection*{Creating knowledge and learning map visualizations}\label{subsec:knowledge-maps}

An important feature of our approach is that, given a trained text embedding
model and participants' quiz performance on each question, we can estimate
their knowledge about \textit{any} content expressible by the embedding
model---not solely the content explicitly probed by the quiz questions, or even
appearing in the lectures. To visualize these estimates
(Fig.~\ref{fig:knowledge-maps}, Supp.
Figs.~\individualKnowledgeMapsA,~\individualKnowledgeMapsB,~\individualKnowledgeMapsC,~\individualLearningMapsA,
and~\individualLearningMapsB), we used Uniform Manifold Approximation and
Projection (UMAP; \citealp{McInEtal18a, McInEtal18b}) to construct a 2D
projection of the text embedding space. Sampling the original 100-dimensional
space at high resolution to obtain an adequate set of topic vectors spanning
the embedding space would be computationally intractable. However, sampling a
2D grid is trivial.

At a high level, the UMAP algorithm obtains low-dimensional embeddings by
minimizing the cross-entropy between the pairwise (clustered) distances between
the observations in their original (e.g., 100-dimensional) space and the
pairwise (clustered) distances in the low-dimensional embedding space (in our
approach, the embedding space is 2D). In our implementation, pairwise distances
in the original high-dimensional space were defined as 1 minus the correlation
between each pair of coordinates, and pairwise distances in the low-dimensional
embedding space were defined as the Euclidean distance between each pair of
coordinates.

In our application, all of the coordinates we embedded were topic vectors,
whose elements are always non-negative and sum to one. Although UMAP is an
invertible transformation at the embedding locations of the original data,
other locations in the embedding space will not necessarily follow the same
implicit ``rules'' as the original high-dimensional data. For example,
inverting an arbitrary coordinate in the embedding space might result in
negative-valued vectors, which are incompatible with the topic modeling
framework. To protect against this issue, we log-transformed the topic vectors
prior to embedding them in the 2D space. When we inverted the embedded vectors
(e.g., to estimate topic vectors or word clouds, as in
Fig.~\ref{fig:knowledge-maps}C), we passed the inverted (log-transformed)
values through the exponential function to obtain a vector of non-negative
values, and normalized them to sum to one.

After embedding both lectures' topic trajectories and the topic vectors of
every question, we defined a rectangle enclosing the 2D projections of the
lectures' and quizzes' embeddings. We then sampled points from a regular $100
\times 100$ grid of coordinates that evenly tiled this enclosing rectangle. We
sought to estimate participants' knowledge (and learning, i.e., changes in
knowledge) at each of the resulting 10,000 coordinates.

To generate our estimates, we placed a set of 39 radial basis functions (RBFs)
throughout the embedding space, centered on the 2D projections for each
question (i.e., we included one RBF for each question). At coordinate $x$, the
value of an RBF centered on a question's coordinate $\mu$, is given by:
\begin{equation}
    \mathrm{RBF}(x, \mu, \lambda) = \exp\left\{-\frac{||x - \mu||^2}{\lambda}\right\}.
    \label{eqn:rbf}
\end{equation}
The $\lambda$ term in the RBF equation controls the ``smoothness'' of the
function, where larger values of $\lambda$ result in smoother maps. In our
implementation we used $\lambda = 50$.  Next, we estimated the ``knowledge''
at each coordinate, $x$, using:
\begin{equation}
    \hat{k}(x) = \frac{\sum_{i \in \mathrm{correct}} \mathrm{RBF}(x, q_i, \lambda)}{\sum_{j = 1}^N \mathrm{RBF}(x, q_j, \lambda)}.
    \label{eqn:rbf-knowledge}
\end{equation}
Intuitively, Equation~\ref{eqn:rbf-knowledge} computes the weighted proportion of
correctly answered questions, where the weights are given by how nearby (in the 2D space)
each question is to the $x$.  We also defined \textit{learning maps} as the coordinate-by-coordinate
differences between any pair of knowledge maps.  Intuitively, learning maps reflect the \textit{change}
in knowledge across two maps.

\section*{Author contributions}

Conceptualization: PCF, ACH, and JRM. Methodology: PCF, ACH, and JRM. Software:
PCF. Validation: PCF. Formal analysis: PCF. Resources: PCF, ACH, and JRM. Data
curation: PCF. Writing (original draft): JRM. Writing (review and editing):
PCF, ACH, and JRM. Visualization: PCF and JRM. Supervision: JRM. Project
administration: PCF. Funding acquisition: JRM.

\section*{Data and code availability}

All of the data analyzed in this manuscript, along with all of the code for
running our experiment and carrying out the analyses may be found at
https://github.com/Con\-text\-Lab/eff\-ic\-ient-learn\-ing-khan.

\section*{Acknowledgements}

We acknowledge useful discussions, assistance in setting up an earlier
(unpublished) version of this study, and assistance with data collection
efforts from Will Baxley, Max Bluestone, Daniel Carstensen, Kunal Jha, Caroline
Lee, Lucy Owen, Xinming Xu, and Kirsten Ziman. Our work was supported in part
by NSF CAREER Award Number 2145172 to JRM. The content is solely the
responsibility of the authors and does not necessarily represent the official
views of our supporting organizations. The funders had no role in study design,
data collection and analysis, decision to publish, or preparation of the
manuscript.

\bibliographystyle{apa}
\bibliography{CDL-bibliography/cdl}
\end{document}